### 2020-01-27 exp.tex
Starting .tex and MCS
little back up before merging

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath} % Advanced math typesetting
\usepackage[utf8]{inputenc} % Unicode support (Umlauts etc.)
\usepackage[ngerman]{babel} % Change hyphenation rules
\usepackage{hyperref} % Add a link to your document
\usepackage{graphicx} % Add pictures to your document
\usepackage{listings} % Source code formatting and highlighting
\usepackage{bookmark}
\usepackage{natbib}

\usepackage{multicol}
\usepackage{fancyhdr}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{Overleaf}
\fancyhead[RE,LO]{Guides and tutorials}
\fancyfoot[CE,CO]{\leftmark}
\fancyfoot[LE,RO]{\thepage}

\title{Imm algorithm implementation for tracking}
\author{
Paiola Lorenzo 000000

Zanolli Erik 198852}
\date{January 2020}

\begin{document}

\maketitle

\section*{Introduction}
\begin{flushleft}
The porpouse of this document is the analysis and performance evaluation of an IMM algorithm's implementation in
a sensor grid tracking an object. The object can move with two different model, random walk and unicycle, governed by
 a Markov chain. The goal is evaluate the best trade-off between error on estimated position and real position and number of 
 messagesses involved in tracking
\end{flushleft}

\begin{multicols}{2}
    \section*{Backgound}
    \begin{flushleft}
        Tracking a object with a changing dynamic require propers algorithms and filters. Retriving physical data with
        observations it's not enough, models achiving the best-fit to our data is needed in order to make a prediction and estimate 
        with reduced uncertainty. Here comes to play a major role the IMM algorithm (Interacting Multiple Models): The data
        from sensor are combined with different dynamical models through the algorithm... 
        \subsection*{Kalman Filter and IMM explained}
        \subsection*{m}
    \end{flushleft}
    \section*{Simulation model implementation}
    Introduction
    \subsection*{Filter}
    \subsection*{Sensor}

    \section*{Result}
    Introduction
    \subsection*{}
\end{multicols}
\end{document}